Profiles of Wind Direction and Speed in Turbulent Ekman Flow

### 2021-09-09 write/manuscript.tex
fixed problems in velocity profile (divergence at z=0); added section 3 to manuscript

\begin{abstract}
  The profiles of wind speed and direction in turbulent Ekman flow
  are formulated based on asymptotic theory and data from direct numerical simulation. 
  %
  The profile of the stream-wise component follows the classical viscous, logarithmic and wake scaling.
  %
  In the outer layer, the velocity component profiles can be described by an Ekman-spiral with adapted
  boundary conditions that result in a reduction of the spiral-like rotation. 
  %
  The span-wise component poses a conceptual challenge to the channel-flow analogy
  in the context of asymptotic matching; it exhibits a mixed scaling in the surface layer, but follows
  outer scaling for most of the outer layer.
  %
\end{abstract}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\section{Introduction}

@@

$Oxyz$, where an approximate alignment (plus/ minus few degrees centigrade) of the direction $Ox$ with the surface shear stress is achieved.

@@

(i)  exactly aligned with the surface shear stress and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and

@@

We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity. 

@@

(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,0,G_3)$ is the geostrophic wind vector),

@@

We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that

@@

  \RE=\frac{G \LR }{\nu}. 

@@

@@

geostrophic wind (with $g_{1}^2+g_{3}^2=1$ by construction).

@@

These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxz$, a no-slip--no-penetration boundary at $y=0$ and
a no-penetration, free-slip boundary at $y=L_y$.

@@

for the pressure field is soled to machine accuracy using a combined spectral/compact approach as

@@

We extend an existing set of simulations for $\RE\in\{125\,000, 281\,250,\ 500\,000\}$ \citep[gray shading; cf.][]{ansorge:BM2014,ansorge:JFM2016} 
by new simulations at higher Reynolds numbers up to $1.28\times10^6$.

@@

\todo{(new and old) Simulations}

@@

  \caption{\label{tab:sim-setup} Direct numerical simulation datasets used in this work}

@@

    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\ 
    \bottomrule 
  \end{tabular} 
\end{table}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview of the new simulations at $\RE_\tau$ up to 3000}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Uniqueness and necessity of new simulations
%
\par
%
PARAGRAPH ON BULK PARAMETERS AND SCALING (cf. \ref{tab:bulk_results})
\begin{table}
  \caption{Bulk characterization of the simulations \label{tab:bulk_results}}
  \begin{tabular}{r r r r r r r }
    \toprule
    Re & $\delta^+=\RE_\tau$& $u_\star$ & $\alpha$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
      \midrule 
      \rowcolor{gray!30} 
      500 & 479 & 0.0619& 25.5 & 0.66& 0.88& 1.31 \\
      \rowcolor{gray!30} 
      750 & 886 & 0.0561& 21.0 & 0.65& 0.90& 1.34 \\
      \rowcolor{gray!30}
      1000& 1403& 0.0530& 18.8 & 0.62& 0.92& 1.30 \\
      \midrule
      1300& 2122& 0.0501& 17.9 & 0.59& 0.85& 1.24 \\
      1600& 2978& 0.0482& 17.2 & 0.61& 0.91& 1.21 \\ 
      \bottomrule
    \end{tabular}
\end{table} 
%
\par
%
PARAGRAPH ON VELOCITY AND STRESS PROFILES cf. \ref{fig:stresses}
\begin{figure}
  \includegraphics[width=0.66\textwidth]{../plot/eddy_viscosity.pdf}
  \caption{Reynolds stresses and eddy viscosity across the Ekman layer plotted in ineer scaling for
    various Reynolds number
    \label{fig:stresses}} 
\end{figure} 
%
\par
%
PARAGRAPH ON TURBULENCE STRUCTURE (2D SLICES, SPECTRA?) cf. Fig.~\ref{fig:slices}

\begin{figure}
  \centerline{\includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}
  \centerline{\includegraphics[width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
  \caption{horizontal slices of turbulence kinetic energy in the Buffer layer (i=10), Logarithmich Layer (i=100) and outer lzayer (i=400); coloring between percentiles 4 and 96 of the respective image. Lower panel:streamwise--vertical intersect through the domain} 
  \label{fig:slices} }

@@

\end{itemize}
\begin{figure}
  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
  \caption{Variation of geostrophic drag, $Z_\star$, and surface veering, $\alpha$, with Reynolds number according to
    the theory by Spalart et al. (1989) and as estimated from DNS data}
  \label{fig:drag_law}
\end{figure}
% 
\subsection{Drag-Law}
A drag-law for Ekman flow determines two key parameters of the flow as a function of Reynolds number alone.
%
These are, first, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
second, the direction of surface shear stress, $\alpha$, also termed wind veering. 
%
A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
engineering context, and it confronts us with a situation where the most appropriate coordinate system for analysis
(namely that aligned with the surface shear stress) is a priori unknown. 
%
We compare our DNS data against a semi-empirical drag-law based on integral consideration \cite{spalart:JFM1989}
and find, as demonstrated in previous work \cite{ansorge:BM2014}, excellent agreement in the range
 $400<\mathbf{Re}<1600$, representing a factor of 16 in variation of viscosity. 
%
\par
%
We also find that the solution of the transient equation involved in estimation of $u_\star$ for a given Reynolds number
$Re_D$ is approximated reasonably by the formulation
\begin{equation}
  u_\star= 4\log(\mathbf{Re}_D))-8
\end{equation}
which quantifies the 'weak' dependence of $u_\star$ on the Reynolds number as an approximately logarithmic one, at least
for problems with a scale separation on the order that is relevant to geophysical problems ($Re_D<10^{8}$).  
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Profile in the outer layer}
%
The outer layer of Ekman flow is characterized by a turning of the wind velocity and the super-geostrophic maximum
that is sustained by momentum convergence at the inflection point of the velocity profile. 
The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
the boundary layer are well-described by a classic Ekman spiral with adapted boundary conditions and a shift
in reference height.
%
\begin{figure}
  \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
  \caption{} 
\end{figure}

@@

  (U,W)^{\alpha+}  = (z^+,0)  

@@

In fact, this law of the wall has no degree of freedom given the drag, i.e. once $u_\star$ and $\alpha$ are defined. 

@@

  U_\text{inner}^{\alpha}^+ = \frac{ y^+ + \gamma_4 (y^+)^{4} + \gamma_6 (y^+)^{6} } { 1 + \gamma_6 /u_{ref} (y^+ )^6}, 

@@

  U_\text{log}^{\alpha}^+ = \frac{1}{\kappa} \log y^+ + C 

@@

  \caption{Shear-aligned profiles of velocity components $U^{\alpha+}$ in inner (left) and outer (right) units.} 

@@

component, the veering requires a value of $V_{top}=U_G\sin\alpha$ in the free stream (and thus also at the top of the boundary layer if we assume

@@

   \left.\frac{\partial V^G}{\partial U^G}\right|_{z=0} &=& \sin\alpha

@@

Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^\alpha-G^\alpha$

@@

  V^\alpha-G^\alpha = f^\alpha_V(z^-).  

@@

Interpreting the shear-aligned spanwise velocity deficit $f_{V}^{\alpha}(z^-)$ as 

@@

This cannot be the case for the variation of $f_{V}^{\alpha}(z^-)$ across the boundary layer (i.e. between $0<z^-<1$)

@@

at the same time, the weak dependence of $\alpha$ and $u_\star$ on $Re$ implies this range depends on $Re$.

>>>>

\begin{abstract}
  The profiles of wind speed and direction in turbulent Ekman flow
  are formulated based on asymptotic theory and data from direct numerical simulation. 
  %
  The profile of the stream-wise component follows the classical viscous, logarithmic and wake scaling.
  %
  In the outer layer, the velocity component profiles can be described by an Ekman-spiral with adapted
  boundary conditions that result in a reduction of the spiral-like rotation. 
  %
  The span-wise component poses a conceptual challenge to the channel-flow analogy
  in the context of asymptotic matching; it exhibits a mixed scaling in the surface layer, but follows
  outer scaling for most of the outer layer.
  %
  Viscous stress scales universally across the boundary layer in inner units while the total stress
  becomes universal as a function of outer height.
  %
  This implies a mixed scaling for the turbulent stress and eddy viscosity across the inner layer
  and convergence to a universal scaling as function of the outer height across the outer layer
  for increasing scale separation vide Reynolds numbers. 
\end{abstract}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\section{Introduction}

@@

$Oxyz$, where an approximate alignment (plus/ minus few degrees) of the direction $Ox$ with the surface shear stress is achieved.

@@

(i)  exactly aligned with the surface shear stress
  \underline{\tau}_\mathrm{sfc} = \left(\begin{array}{c} \tau_x\\ \tau_y\\\tau_z\end{array}\right) = - \nu \left( \frac{\partial U}{\partial z} \hat{e}_x + \frac{\partial V}{\partial z} \hat{e}_y\right)
and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and

@@

We denote the normalized modulus of the surface friction by
  u_\star = \sqrt{ || - \underline{\tau}_\mathrm{sfc} || } 
  \label{eqn:ustar_def}
\end{equation} and let $Z_\star=G/u_\star$;
the surface veering angle $\alpha_\star$ is the angle between $\underline{\tau}$ and the geostrophic wind 
  \alpha_\star = \sphericalangle \left(\underline{G}, \underline{\tau}_\mathrm{sfc} \right). 
Analogously, we denote the height-local veering of the wind
$\alpha(z)= \sphericalangle \left( \underline{G},\underline{U}(z)\right)$. 

@@

(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,G_2,0)$ is the geostrophic wind vector),

@@

We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that the Reynolds number governing the problem reads as

@@

  \RE_\Lambda=\frac{G \LR }{\nu}. 

@@

The scales used in defining $\RE_\Lambda$ loose there relevance in describing the turbulent state of the flow. The friction Reynolds number
  \RE_\tau = \frac{u_\star\delta}{\nu} = \frac{\RE_\Lambda}{Z_\star^2}, 
where $\delta=u_\star/f$, gives a measure for the turbulence scale separation that is comparable to other
wall-bounded flow geometries such as rotating or non-rotating channel, pipe, or Couette flow.

@@

geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction).

@@

These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxy$, a no-slip--no-penetration boundary at $z=0$, and
a no-penetration, free-slip boundary at $z=L_z$.

@@

for the pressure field is solved to machine accuracy using a combined spectral/compact approach as

@@

We extend an existing set of simulations for $\RE_\Lambda\in\{125\,000; 281\,250;\ 500\,000\}$ \citep[gray shading; cf.][]{ansorge:BM2014,ansorge:JFM2016} 
by new simulations at higher Reynolds numbers up to $1.28\times10^6$ with a horizontal domain extent up to $3.3\times 10^4$ viscous units. 

@@

@@

  \caption{Direct numerical simulation datasets used in this
    work \label{tab:sim-setup}}

@@

    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\ 
    \bottomrule 
  \end{tabular} 
\end{table}
\section{Scaling behavior of the flow for $\RE_\tau$ up to 3000}
The generalization of profiles to arbitrary Reynolds numbers 
requires sufficient scale separation in the simulations to not only quantify the effect of
the Reynolds number on low-order flow statistics, but also assess the corresponding rate-of-change
to eventually allow for an extrapolation of the findings.
While the simulations available give confidence in a first-order representation of the turbulent
problem, the estimation of higher-order effects such as the dependency of the Reynolds number requires
a broader scale separation that is made available by the two new simulations for 
(cf. Tab.~\ref{tab:sim-setup}). 
%%% We thus require sufficient scale separation; while we are certain that even the lower Reynolds numbers
%%% are well within the turbulent regime, the range of separation required to estimate $\RE$-effects should
%%% be about one magnitude at the least, i.e. we anticipate that data for $\mathcal{O}(\RE)=10^6$ would 
%%% allow us to extrapolate findings. 
Data at such scale separation has been obtained previously \citep[cf.][]{spalart:PF2008,spalart:PF2009}, 
but we also need high confidence with respect to the convergence of simulation data for slow oscillations 
and with respect to sampling convergence, which translates to two further requirements on the data: 
First, data should be free of artifacts from long-term oscillations across the vertical
extent of the domain--primarily, simulations should be free of effects originating from the
inertial oscillation; this is achieved here by replacing the mean value of the three-dimensional
velocity fields by the time mean over a whole inertial oscillation. 
Second, high accuracy is also needed in terms of the statistical convergence of averages,
bulk measures and large-scale structures; this requires a domain size $L_{x} > \mathcal{O}(\delta_{95})$. 
We use here $L_{x}=L_{y}=1.08\Lambda$ for cases with $\RE \le 1000$ and
$L_{x}=L_{y}=0.54\Lambda$ for $\RE \ge 1300$ which corresponds to $L_{x}/\delta_{95}\approx 25$ for $Re_D=500$
and $L_{x}/\delta_{95} \approx 18$ for $Re_D=1300$.
Bulk parameters of the simulations are given in Tab.~\ref{tab:bulk_results}.
The surface stress is characterized by $u_\star$ and $\alpha_\star$ in relation to the geostrophic wind vector
and discussed in more detail as the drag law below in Sec.~\ref{sec:drag-law}
(we find the expected slight decrease of $u_\star$ and $\alpha_\star$ with increasing $\RE$).
The boundary-layer height estimated from the 95\% stress reduction, $\delta_{95}$, is around $0.6\delta$ to $0.66\delta$. 
Interestingly, the integrated TKE $\int_{0}^{\delta} e\mathrm{d}z$ stays constant when normalized by the friction velocity $u_\star$
while the integrated dissipation $\int_{0}^{\delta} \epsilon\mathrm{d}z$ exhibits inviscid scaling when normalized
by the magnitude $G$ of the geostrophic wind. 
(Alternatively normalized TKE and dissipation,
$f G^{-3} \int e \mathrm{d} z$ and $u_\star^{-3} \int \epsilon\mathrm{d}z$, exhibit substantial
dependence on $\RE$ for the variation of $u_s\tar$.)
This indicates that the bulk dissipation is governed by the forcing $G$--irrespective of $\RE$. 
Changes in $\RE$, however, affect the level and organization of turbulence, and the parameter
representing this dependency is the friction velocity $u_\star$ which describes the turbulence
production processes in the surface layer, in particular in the buffer layer. 
  \caption{Bulk characterization of the simulations for different Reynolds numbers 
    Viscous Reynolds number $\RE$,
    friction Reynolds number  $\RE_\tau$,
    friction velocity $u_\star$,
    surface veering angle $\alpha_\star$,
    normalized boundary layer depth $\delta_{95}/\delta$
    inner normalization of vertically integrated TKE,
    outer normalization of vertically integrated dissipation.
    \label{tab:bulk_results}}
  \begin{tabular}{r r r r r r r }
    Re & $\delta^+=\RE_\tau$& $u_\star$ & $\alpha_\star$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
      500 & 479 & 0.0619& 25.5 & 0.66& 0.88& 1.31 \\
      750 & 886 & 0.0561& 21.0 & 0.65& 0.90& 1.34 \\
      1000& 1403& 0.0530& 18.8 & 0.62& 0.92& 1.30 \\
      1300& 2122& 0.0501& 17.9 & 0.59& 0.85& 1.24 \\
      1600& 2978& 0.0482& 17.2 & 0.61& 0.91& 1.21 \\ 
      \bottomrule
    \end{tabular}
\end{table} 
Velocity profiles in inner and units (Fig.~\ref{fig:profiles}a,b) are in accordance
with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:
The profile of the shear-aligned streamwise velocity component are well-collapsed for the Reynolds
numbers investigated below $z^-\approx 0.15$ (circles in Fig.~\ref{fig:profiles}a); the logarithmic
law is appropriate for $50 < z^+ < 0.15 Re_\tau$, where $z^+ = Re_\tau z^-$.
While the profiles $U^{\alpha_\star+}(z^+)$ diverge between different $\Re$ beyond $z^-=0.15$,
the corresponding profiles $U^{\alpha_\star}(z^-)/u_\star$ agree closely,
irrespective of $\RE$ which illustrates the
inner--outer scale-duality in this external flow. 
It is noteworthy that also in the outer layer of the flow, $u_\star$--and not the magnitude the
geostrophic wind speed $G$--is the governing velocity scale for inviscid, i.e. $\RE$-independent,
normalization. \todo{WHY? cf. TKE/Dissipation normalization above}
No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$;
When normalized in outer units, the profiles of spanwise velocity $V^{\alpha_\star}(z^-)/u_\star$
agree well beyond $z^-\approx 0.3$ whereas the profiles of the streamwise component collapse also
within the overlap layer that extends much closer down to the surface
(circles in Fig.~\ref{fig:profiles}b).
While low-$\RE$ effects appear to be present for $\mathcal{O}(\RE)<10^3$, the spanwise component
appears to converge to an $\RE$-independent limit within the range of scale separation considered
here.
  \begin{flushleft}
    \textbf{a)}\hspace{0.47\textwidth} \textbf{b)}\\
    \includegraphics[width=\textwidth]{../plot/uv_innerouter.pdf}\\
    \caption{
      \textbf{a)} Shear-aligned velocity profiles in inner units;
      \textbf{b)} Shear-aligned velocity deficit in outer units; circles mark the
      height $z^-=0.15$; dashed line gives the best logarithmic fit to the data. 
      \label{fig:profiles}
    }
  \end{flushleft}

\begin{figure}
  \begin{flushleft} 
    \textbf{a)}\hspace{0.47\textwidth} \textbf{b)}\\
    \includegraphics[width=\textwidth]{../plot/stresses.pdf}
    \textbf{c)}\hspace{0.47\textwidth} \textbf{d)}\\    
    \includegraphics[width=\textwidth]{../plot/eddy_viscosity.pdf}\\
  \end{flushleft} 
  \caption{
    \textbf{a-b)} Profiles of
    the turbulent stress $S_\mathrm{turb}^+$ (solid), 
    the viscous stress  $S_\mathrm{visc}^+$ (dotted), 
    and the total stress, $S^+=S_\mathrm{visc}^++S_\mathrm{turb}^+$ (dashed)
    as a function of inner height \textbf{(a)} and outer height \textbf{(b)}. 
    \textbf{c-d)} Normalized eddy viscosity $\nu_E$ (solid) plotted versus
    inner \textbf{(c)} and outer height \textbf{(d)}.
    Different colors are for different Reynolds numbers (cf. Tab.~\ref{tab:sim-setup}).
    Circles in \textbf{(a)} and \textbf{(b)} denote the height
    $z^-=0.15$, diamonds in \textbf{(c)} and \textbf{(d)} are for $z^+=50$
    as in Fig.~\ref{fig:profiles}.
    \label{fig:stresses}}  
\end{figure} 
The normalized viscous stress
\begin{subequations}
\begin{equation}
  S_\mathrm{visc}= \nu \sqrt\left(\frac{\p U}{\p z}\right)^2  + \left(\frac{\p V}{\p z}\right)^2}
\end{equation}
exhibits universal scaling when considered as $S_\mathrm{visc}^+(z^+)$ (Fig.~\ref{fig:stresses}c); this normalization is also appropriate in the outer layer where the viscous stress
is, however, small. 
%
Small deviations from the universal profile are observed for the smallest Reynolds number $\RE=500$; 
we attribute these to low-$\RE$ effects in the only transitionally turbulent flow ($\RE_\tau=479$). 
%
In contrast to the viscous stress, the total stress follows outer normalization, i.e.
$S^+(z^-)$ is universal.
%
This implies a mixed scaling for the turbulent stress, 
\begin{equation}
  S_\mathrm{turb} = \sqrt{\overline{uw}^2 + \overline{vw}^2}. 
\end{equation} 
Indeed, $S_\mathrm{turb}$ only follows inner normalization below $z^+\lesssim 20$
(where the turbulent contribution is negligible).
%
In the outer layer, where $S_\mathrm{visc}\rightarrow 0$, $S_\mathrm{turb}^+$
follows outer normalization for $z^-\gtrsim0.15$---with increasing accuracy for larger $\RE$ and larger distance from
the surface. 
%
In the overlap region, i.e. for $z^+>20$ and $z^-<0.15$, the mixed scaling for the
turbulent stress can be expressed as
\begin{equation}
  S_\mathrm{turb}^+(z^+,\RE_\tau) = S^+(z^-) - S_\mathrm{visc}^+(z^+), 
\end{equation}
where $z^-=z^+/\RE_\tau$. 
%
\par
%
The Eddy viscosity plays a crucial part when modelling profiles and the vertical
transport in turbulent flow. 
%
In analogy to the Fick-law for molecular diffusion, the eddy diffusivity is
the effective diffusivity that yields the turbulent transport $S_\mathrm{turb}$ based
on the strain rate.
%
For the symmetries in the flow (horizontal homoegeneity, and $W=0$), it is
%
\begin{equation}
  \nu_{E}= \frac{S_\mathrm{turb}}{\sqrt{\left(\frac{\p U}{\p z} \right)^2 + \left(\frac{\p V}{\p z}\right)^2}} = \nu \frac{S_\mathrm{turb}}{S_\mathrm{visc}}.  
\end{equation}
%
The inner normalization of $\nu_E$ s

\begin{eqnarray}
 \nu_E^+  \frac{Z_{\star}^{2}}{\delta+} &=& \frac{\nu_E^+}{\nu}
\end{eqnarray} 
\par
%
\todo{PARAGRAPH ON TURBULENCE STRUCTURE (2D SLICES, SPECTRA?) cf. Fig.~\ref{fig:slices}}

\begin{figure}
  \centerline{
    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
    \includegraphics[trim=1152 0 0 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}
  \centerline{
    \includegraphics[trim=1152 0 0 0, clip, width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
  \caption{horizontal slices of turbulence kinetic energy in the
    Buffer layer (i=10),
    logarithmic Layer (i=100), and
    outer layer (i=400);
    coloring between percentiles 4 and 96 of the respective image. Lower panel:streamwise--vertical intersect through the domain 
    \label{fig:slices}}

@@

\end{itemize}
\begin{figure}
  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
  \caption{Variation of geostrophic drag, $Z_\star$, and surface veering, $\alpha_\star$, with Reynolds number according to
    the theory by Spalart et al. (1989) and as estimated from DNS data
    \label{fig:drag_law}}
\end{figure}
% 
\subsection{Drag-Law}
\label{sec:drag-law}
A drag-law for Ekman flow determines two key parameters of the flow as a function of Reynolds number alone.
%
These are, first, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
second, the direction of surface shear stress, $\alpha_\star$, also termed wind veering. 
%
A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
engineering context, and it confronts us with a situation where the most appropriate coordinate system for analysis
(namely that aligned with the surface shear stress) is a priori unknown. 
%
We compare our DNS data against a semi-empirical drag-law based on integral consideration \cite{spalart:JFM1989}
and find, as demonstrated in previous work \cite{ansorge:BM2014}, excellent agreement in the range
 $400<\mathbf{Re}<1600$, representing a factor of 16 in variation of viscosity. 
%
\par
%
We also find that the solution of the transient equation involved in estimation of $u_\star$ for a given Reynolds number
$Re_D$ is approximated reasonably by the formulation
\begin{equation}
  u_\star= 4\log(\mathbf{Re}_D))-8
\end{equation}
which quantifies the 'weak' dependence of $u_\star$ on the Reynolds number as an approximately logarithmic one, at least
for problems with a scale separation on the order that is relevant to geophysical problems ($Re_D<10^{8}$).  
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Profile in the outer layer}
%
The outer layer of Ekman flow is characterized by a turning of the wind velocity and the super-geostrophic maximum
that is sustained by momentum convergence at the inflection point of the velocity profile. 
The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
the boundary layer are well-described by a classic Ekman spiral with adapted boundary conditions and a shift
in reference height.
%
\begin{figure}
  \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
  \caption{} 
\end{figure}

@@

  (U,W)^{\alpha_\star+}  = (z^+,0)  

@@

In fact, this law of the wall has no degree of freedom given the drag, i.e. once $u_\star$ and $\alpha_\star$ are defined. 

@@

  U_\text{inner}^{\alpha_\star}^+ = \frac{ y^+ + \gamma_4 (y^+)^{4} + \gamma_6 (y^+)^{6} } { 1 + \gamma_6 /u_{ref} (y^+ )^6}, 

@@

  U_\text{log}^{\alpha_\star}^+ = \frac{1}{\kappa} \log y^+ + C 

@@

  \caption{Shear-aligned profiles of velocity components $U^{\alpha_\star+}$ in inner (left) and outer (right) units.} 

@@

component, the veering requires a value of $V_{top}=U_G\sin\alpha_\star$ in the free stream (and thus also at the top of the boundary layer if we assume

@@

   \left.\frac{\partial V^G}{\partial U^G}\right|_{z=0} &=& \sin\alpha_\star

@@

Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^\alpha_star-G^\alpha_\star$

@@

  V^\alpha_\star-G^\alpha_\star = f^\alpha_\star_V(z^-).  

@@

Interpreting the shear-aligned spanwise velocity deficit $f_{V}^{\alpha_\star}(z^-)$ as 

@@

This cannot be the case for the variation of $f_{V}^{\alpha_\star}(z^-)$ across the boundary layer (i.e. between $0<z^-<1$)

@@

at the same time, the weak dependence of $\alpha_\star$ and $u_\star$ on $Re$ implies this range depends on $Re$.
